---
title: "Zika Analysis"
author: "Justin Lessler"
date: "Monday, February 22, 2016"
output: word_document
---

Pooled analysis of data form the Zika systematic review.

First load the data and figure out how big our data set is for each purpose.

```{r}
zika <- read.csv("ZikaLitReviewData.csv")

#How many people do we have?
length(unique(zika$UID))

#Now lets divide data sets by observation type.

#Incubation period
zika.inc <- zika[!is.na(zika$SL),]
nrow(zika.inc)

#Period of viral shedding
zika.vshed <- zika[!is.na(zika$VL),]
nrow(zika.vshed)
length(unique(zika.vshed$UID))

#Time to seropositiveity
zika.sero <- zika[!is.na(zika$AL),]
nrow(zika.sero)
length(unique(zika.sero$UID))

```

So we have data with which to estimate the incubation period from 20 people; 40 observations from 17 people to estimate the length of viral shedding; and 109 observations from 21 people to estimate the time to seropositivity.


#Incubation Period

Basic incubation period analysis assuming a log-normal incubation period and using a bootstrap method for calculating confidence intervals. 

```{r, results='hide'}
library(coarseDataTools)

#make the data to pass in to dic.fit
inc.dat <- zika.inc[,c("EL","ER","SL","SR")]
inc.dat$type <- 0

#fit it...use bootstrap interstions for now. MCMC better?
#inc.fit <- dic.fit(inc.dat,dist="L",n.boots=1000,
#                   ptiles = c(0.05, 0.25, 0.5, 0.75, 0.95))


inc.fit.mcmc <- dic.fit.mcmc(inc.dat, n.samples = 2000,
                              ptiles = c(0.05, 0.25, 0.5, 0.75, 0.95))



```

```{r}
plot(inc.fit.mcmc, xlim=c(0,20), plot.n.samps=2000)
inc.fit.mcmc
```


We estimate the mean incubation period of Zika virus to be 5.7 days (95% CI: 3.9-7.9). The dispersion is 1.60 (95% CI: 1.31-2.42). Hence 5% of cases will develop symptoms by 2.6 days after infection (95% CI: 1.1-4.0), 25% by 4.1 days (95% CI: 2.4-5.7), 75% by 7.8 days (95% CI: 5.6-12.1) and 95% by 12.4 days (95% CI: 7.9-26.0).



#Length of viral shedding

```{r}

#first pass, restrict to no blood
zika.vshed.nb <- zika.vshed[which(zika.vshed$SAMPLE.TYPE %in%c("serum","blood")),]

#Go through and find the start of the latest period in whihc
#someone tested seropositive and the end of the latest subsequent
#period where they did not....these bound seroconversion
for (uid in unique(zika.vshed.nb$UID)) {
  inds <- which(zika.vshed.nb$UID==uid)
  tmp <- zika.vshed.nb[inds,] 
  low.time <- max(tmp$VL[tmp$SHEDDING==1])
  
  #no earliest point...go to earliest time of infection
  if (low.time<tmp$EL[1]) {low.time <- tmp$EL[1]}
  
  high.time <- min(tmp$VR[tmp$SHEDDING==0 &  tmp$VR>low.time])
  ##print(tmp) ###DEBUG
  ##cat(uid, low.time, high.time,"\n") ###DEBUG
  
  zika.vshed.nb$VCL[inds] <- low.time
  zika.vshed.nb$VCR[inds] <- high.time
  }


  #make the anlytic dataset
  vshed.nb.anal <- zika.vshed.nb[,c("UID","EL","ER","VCL","VCR")]
  vshed.nb.anal$type <- 0
  vshed.nb.anal <- unique(vshed.nb.anal) #one row per observation

  ###REPLACED BY STAN ANALYSIS
  #colnames(vshed.nb.anal) <- c("UID","EL","ER","SL","SR","type")

  #change the infinite SRs to something really big
  #vshed.nb.anal$SR[is.infinite(vshed.nb.anal$SR)] <- 
  #    vshed.nb.anal$ER[is.infinite(vshed.nb.anal$SR)] + 730

  #now lets fit a gamma distribution to this data using MCMC
  #vshed.fit <- dic.fit(vshed.nb.anal, dist = "G", n.boots = 10000)
  #vshed.fit.mcmc <- dic.fit.mcmc(vshed.nb.anal, dist = "L",
  #                               burnin = 300,n.samples=500, optim.method="L-BFGS-B")

```


#Combined stan analysis Lognormal for all distributions
```{r}
require(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.nb.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720

zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#Replace

#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR)


#initialization version for the lognormal/lognormal stan model
stan.init.LL <- function() {
    rc <-list(E=runif(stan.data$N, stan.data$EL, stan.data$ER),
                lm=runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_lm = runif(1,log(2),log(10)),
                v_lsd = runif(1,.1,2) 
                )
    print(rc)
    return(rc)
    }



#Need at least 20,000 iterations to converge
full.fit <- stan("DistributionFitLL.stan", data=stan.data,
                  init=stan.init.LL,
                  iter=20000, chains=3)


#now we need to calculate the quantiles from the chains and construct our
#data frame
chains <- extract(full.fit)


inc.fit.stanLL <- c(mean(chains$lm), quantile(chains$lm,prob=c(0.025,0.975)))
inc.fit.stanLL <- rbind(inc.fit.stanLL,
        c(mean(chains$lsd), quantile(chains$lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$lm, chains$lsd)
  inc.fit.stanLL <- rbind(inc.fit.stanLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(inc.fit.stanLL) <- c("est","CIlow","CIhigh")
rownames(inc.fit.stanLL) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")



vshed.fit.stanLL <- c(mean(chains$v_lm), quantile(chains$v_lm,prob=c(0.025,0.975)))
vshed.fit.stanLL <- rbind(vshed.fit.stanLL,
        c(mean(chains$v_lsd), quantile(chains$v_lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$v_lm, chains$v_lsd)
  vshed.fit.stanLL <- rbind(vshed.fit.stanLL,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(vshed.fit.stanLL) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.stanLL) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")


```



#Combined stan analysis Lognomal for incubation period, Gamma others

This is not working at this point.
```{r}
require(rstan)


#Create a dataset with everythign lined up for passing into STAN
zika.anal <- zika.inc[,c("UID","EL","ER","SL","SR")]
zika.anal <- merge(zika.anal, vshed.nb.anal, all=TRUE)

zika.anal$SL[which(is.na(zika.anal$SL))] <- 
  zika.anal$EL[which(is.na(zika.anal$SL))]
zika.anal$SR[which(is.na(zika.anal$SR))] <- 
   zika.anal$ER[which(is.na(zika.anal$SR))] + 100

zika.anal$VCL[which(is.na(zika.anal$VCL))] <-
  zika.anal$EL[which(is.na(zika.anal$VCL))]
zika.anal$VCR[which(is.na(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.na(zika.anal$VCR))] + 720

zika.anal$VCR[which(is.infinite(zika.anal$VCR))] <- 
  zika.anal$ER[which(is.infinite(zika.anal$VCR))] + 720

#Replace

#for STAN purposes, it is usefule if all of these variables index 
#from 0
stan.data <- list(N=nrow(zika.anal),
                  EL=zika.anal$EL,
                  ER=zika.anal$ER,
                  SL=zika.anal$SL,
                  SR=zika.anal$SR,
                  VL=zika.anal$VCL,
                  VR=zika.anal$VCR)


#initialization version for the lognormal/lognormal stan model
stan.init.LG <- function() {
    rc <-list(E=runif(stan.data$N, stan.data$EL, stan.data$ER),
                lm=runif(1,log(2),log(10)),
                lsd = runif(1,.1,2),
                v_alpha = runif(1,5,10),
                v_beta= runif(1,.1,1) 
                )
    print(rc)
    return(rc)
    }



#Need at least 20,000 iterations to converge
full.fit <- stan("DistributionFitLG.stan", data=stan.data,
                  init=stan.init.LG,
                  iter=200, chains=3)


#now we need to calculate the quantiles from the chains and construct our
#data frame
chains <- extract(full.fit)


inc.fit.stan <- c(mean(chains$lm), quantile(chains$lm,prob=c(0.025,0.975)))
inc.fit.stan <- rbind(inc.fit.stan,
        c(mean(chains$lsd), quantile(chains$lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$lm, chains$lsd)
  inc.fit.stan <- rbind(inc.fit.stan,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(inc.fit.stan) <- c("est","CIlow","CIhigh")
rownames(inc.fit.stan) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")



vshed.fit.stan <- c(mean(chains$v_lm), quantile(chains$v_lm,prob=c(0.025,0.975)))
vshed.fit.stan <- rbind(vshed.fit.stan,
        c(mean(chains$v_lsd), quantile(chains$v_lsd,prob=c(0.025,0.975))))
for (q in c(0.05, 0.25, 0.5, 0.75, 0.95)) {
  tmp <- qlnorm(q, chains$v_lm, chains$v_lsd)
  vshed.fit.stan <- rbind(vshed.fit.stan,
                        c(mean(tmp), quantile(tmp, prob=c(0.025, 0.975))))
}

colnames(vshed.fit.stan) <- c("est","CIlow","CIhigh")
rownames(vshed.fit.stan) <- c("meanlog",
                            "sdlog",
                            "p5","p25","p50","p75","p95")


```
